---
title: "Final Tech Capacity Figures"
author: "EmmaLi Tsai"
date: "2024-03-06"
output: html_document
---

```{r}
# updated Oct 2024 for updated OPM trends 
library(tidyverse)
library(plotly)
library(htmlwidgets)
library(aws.s3)
library(rvest)

source("./functions/assemble_cube.R")
source("./functions/download_cubes.R")
```

## Grabbing & assembling data cubes from March 2018 - March 2023 
```{r}
# note the folder you'd like to store and pull cubes from: 
cube_folder <- "./data/"
download_cubes(url = "https://www.opm.gov/data/datasets/Index.aspx?tag=FedScope", 
               dest_folder = cube_folder, up_to_year = 2013, 
               month = "March")

# NOTE !! you'll have to manually add employment cubes beyond 2022 due to an 
# error in their HTML tags 
# you can do that easily from the full data list: https://www.opm.gov/data/datasets/Index.aspx

# looking at 10-year trends: 
folder_list <- list.files(cube_folder)
cube_data_test <- assemble_cube(parent_folder = cube_folder, 
                                folder_list = folder_list)
cube_data_10_copy <- cube_data_test
cube_data_10_copy <- cube_data_10_copy %>%
  slice(-1)
# 20 million rows! 

# looking at 5-year trends 
# cube_folder <- "./data/original_manual/"
# cube_data_5yr <- assemble_cube(parent_folder = cube_folder, 
#                                 folder_list = folder_list)
# cube_data_5_copy <- cube_data_5yr
# cube_data_5_copy <- cube_data_5_copy %>%
#   slice(-1)
# wooooow 12 million rows 

# toggling temporal scale: 
cube_data <- cube_data_10_copy
```

## Filtering tidying trends for environmental agencies: 
```{r}
# environmental agencies: 
# DOA (AG), DOC-NOAA (CM-CM54), DOI (IN), EPA (EP)
enviro_agy_codes <- c("AG", "IN", "EP")
enviro <- cube_data %>%
  filter(agy %in% enviro_agy_codes) 
# grabbing NOAA: 
noaa <- cube_data %>%
  filter(agysub == "CM54")
usace <- cube_data %>%
  filter(agysub == "ARCE")
# adding to other environmental agencies: 
enviro <- rbind(enviro, noaa, usace) %>%
  mutate(employment = as.numeric(employment)) %>%
  mutate(salary = as.numeric(salary)) %>%
  mutate(year = as.numeric(year))
enviro$agy <- as.factor(enviro$agy)
levels(enviro$agy) <- c("DOA", "USACE", "DOC-NOAA", "EPA", "DOI")

# tidying string names across the dataframe:   
enviro <- enviro %>%
  separate(agysubt, c("abbr", "name"), "-") %>%
  mutate(name = str_to_title(name), 
         # name = gsub("U.s. ", "", name), 
         occt = str_to_title(occt), 
         los = as.numeric(los), 
         agyt = str_to_title(agyt)) 
# writing 
# write.csv(enviro, "./data/enviro_2013_2023.csv")
# adding to s3: 
# tmp <- tempfile()
# write.csv(enviro, file = paste0(tmp, ".csv"))
# on.exit(unlink(tmp))
# put_object(
#   file = paste0(tmp, ".csv"),
#   object = "/tech-capacity/enviro_staffing.csv",
#   bucket = "tech-team-data",
# )
```

Tidying code for primary figures
```{r}
# reading in data - updated with 2024 data 
enviro <- aws.s3::s3read_using(read.csv, 
                               object = "s3://tech-team-data/tech-capacity/enviro_staffing.csv")

# enviro <- read.csv("./data/enviro_2013_2023.csv")

# in stem: 2210, 1550, 0854, 1515
tech_job <- c("1550", "2210", "1515", "0854")
# computer scientists, IT technology management, operations research, 
# computer engineering

# in stem: "1530" "1529" "1560" "0150" "1370"
data_analyst <- c("1560", "1370", "0150", "1530", "1529")
# data scientists, cartographers, geographers, cartographic technician, 
# geodetic technician, statistics, mathematical statistics 

bureaus <- c("FOREST SERVICE", "NATURAL RESOURCES CONSERVATION SERVICE", 
             "U.S. FISH AND WILDLIFE SERVICE", "GEOLOGICAL SURVEY", 
             "BUREAU OF RECLAMATION", "BUREAU OF LAND MANAGEMENT", 
             "NATIONAL PARK SERVICE", "ENVIRONMENTAL PROTECTION AGENCY", 
             "U.S. ARMY CORPS OF ENGINEERS", "NATIONAL OCEANIC AND ATMOSPHERIC ADMINISTRATION")
bureaus <- str_to_title(bureaus)

# Grabbing stem jobs: 
enviro_tidy <- enviro %>%
  mutate(job_type_broad = case_when(!(stemocc %in% c("XXXX", "****")) ~ "STEM_research", 
                              TRUE ~ "NOT_STEM")) %>%
  mutate(job_type_specific = case_when(stemocc %in% tech_job ~ "STEM_tech",
                              stemocc %in% data_analyst ~ "STEM_analyst", 
                              TRUE ~ job_type_broad)) %>%
  mutate(job_type_broad = case_when(job_type_broad == "STEM_research" ~ "STEM", 
                                    TRUE ~ job_type_broad))
# tidying agency names
enviro_tidy_final <- enviro_tidy %>%
  filter(name %in% bureaus) %>%
  mutate(name = str_to_title(name)) %>%
  mutate(name = gsub("U.s. ", "", name)) %>%
  mutate(agy = case_when(agy == "DOA" ~ "USDA", 
                         agy == "EPA" ~ "EPA", 
                         agy == "USACE" ~ "Army", 
                         agy == "DOC-NOAA" ~ "DOC", 
                         TRUE ~ agy)) %>%
  mutate(agy_name = paste0(agy,  "-", name)) %>%
  mutate(agy_name = 
           case_when(agy_name == "DOC-National Oceanic And Atmospheric Administration" ~ "DOC-NOAA", 
                     agy_name == "Army-Army Corps Of Engineers" ~ "Army Corps of Engineers",
                     TRUE ~ agy_name)) 

# calculating total employees and net change
all_emp <- enviro_tidy_final %>%
  select(-X) %>%
  group_by(agy, agy_name, year) %>%
  summarize(total_employees = sum(employment))

net_change <- all_emp %>%
  filter(year %in% c(2013, 2023)) %>%
  pivot_wider(names_from = year, values_from = total_employees) %>%
  mutate(net_change = `2023` - `2013`)
  
# also wanting to add larger DOI & USDA trends: 
# just_usda_doi <- enviro_tidy %>%
#   mutate(name = str_to_title(name)) %>%
#   mutate(name = gsub("U.s. ", "", name)) %>%
#   mutate(agy = case_when(agy == "DOA" ~ "USDA", 
#                          agy == "EPA" ~ "EPA", 
#                          agy == "USACE" ~ "Army", 
#                          agy == "DOC-NOAA" ~ "DOC", 
#                          TRUE ~ agy)) %>%
#   mutate(agy_name = paste0(agy,  "-", name)) %>%
#   mutate(agy_name = 
#            case_when(agy_name == "DOC-National Oceanic And Atmospheric Administration" ~ "DOC-NOAA", 
#                      agy_name == "Army-Army Corps Of Engineers" ~ "Army Corps of Engineers",
#                      TRUE ~ agy_name)) 
# 
# just_usda_doi <- just_usda_doi %>%
#   select(-X) %>%
#   group_by(agyt, year) %>%
#   summarize(total_employees = sum(employment)) %>%
#   filter(agyt %in% c("Ag-Department Of Agriculture", "In-Department Of The Interior")) %>%
#   mutate(agyt = case_when(
#     agyt == "Ag-Department Of Agriculture" ~ "USDA", 
#     TRUE ~ "DOI")) %>%
#   rename(agy_name = agyt) %>%
  # mutate(agy = agy_name)
```

## Plot One: Total Employment Trends
```{r}
total_emp_plot <- ggplot(all_emp, aes(x = year, y = total_employees, 
                                      color = agy_name)) + 
  geom_point(size = 1.5) + 
  facet_wrap(~agy, scales = "free") + 
  geom_smooth(se = FALSE, size = 1.5) + 

  scale_color_manual(values = c("#606c38", "#2a9d8f",
                                "#fb8500", "#ffb703",
                                "#d62828", "#b5838d", 
                                "#e5989b", "#84a98c", 
                                "#9a8c98", "#4a4e69"), 
                     name = "") + 
  theme_minimal() +
  theme(legend.position = "bottom", 
        axis.text.x = element_text(angle = 90)) + 
  guides(color = guide_legend(ncol=3)) + 
  scale_x_continuous(breaks = seq(min(all_emp$year),
                                  max(all_emp$year),
                                  1)) + 
  labs(x = "Year", y = "Total Employees") + 
  theme(text=element_text(size=11))
total_emp_plot
total_emp_plot <- plotly::ggplotly(total_emp_plot)
```

## Plot two: Tech and STEM Workfroce Breakdown Trends: 
```{r}
# calculating workforce breakdowns: 
stem_breakdown_jobs_b <- enviro_tidy %>%
  filter(name %in% bureaus) %>%
  mutate(name = str_to_title(name)) %>%
  group_by(agy, name, year, job_type_specific) %>%
  summarize(job_specific_employment = sum(employment)) %>%
  mutate(name = gsub("U.s. ", "", name)) %>%
  mutate(agy = case_when(agy == "DOA" ~ "USDA", 
                         agy == "EPA" ~ "EPA", 
                         agy == "USACE" ~ "DoArmy", 
                         agy == "DOC-NOAA" ~ "DOC", 
                         TRUE ~ agy)) %>%
  mutate(agy_name = paste0(agy,  "-", name)) %>%
  mutate(agy_name = case_when(agy_name == "DOC-National Oceanic And Atmospheric Administration" ~ "DOC-NOAA", 
                              TRUE ~ agy_name)) 

stem_jobs_wide <- pivot_wider(stem_breakdown_jobs_b, 
                              names_from = job_type_specific,
                              values_from = job_specific_employment) %>%
  mutate(STEM_tech = replace_na(STEM_tech, 0),
           non_tech = NOT_STEM + STEM_analyst + STEM_research, 
         ratio_tech = STEM_tech/non_tech)

stem_jobs_wide$ratio_tech <- round(stem_jobs_wide$ratio_tech, 4)

# this is the ratio we're after for Oct 2024: 
# write.csv(stem_jobs_wide, "./results/federal_tech_ratio.csv")
# tmp <- tempfile()
# write.csv(stem_jobs_wide, file = paste0(tmp, ".csv"))
# on.exit(unlink(tmp))
# put_object(
#   file = paste0(tmp, ".csv"),
#   object = "/tech-capacity/federal_tech_ratio.csv",
#   bucket = "tech-team-data",
# )

# reorganizing now that we've consolidated STEM and into analyst and 
# research positions: 
stem_jobs_reorg <- stem_jobs_wide %>%
  mutate(STEM = STEM_analyst + STEM_research, 
         total_employees = NOT_STEM + STEM_tech + STEM) %>%
  select(-c("STEM_analyst", "STEM_research", 
            "non_tech", "ratio_tech")) %>%
  rename(Tech = STEM_tech, 
         `Not STEM` = NOT_STEM) %>%
  pivot_longer(`Not STEM`:STEM, names_to = "job_group")  %>%
  mutate(percent_workforce = round((value/total_employees)*100, 3))

stem_jobs_bureau <- stem_jobs_reorg %>%
  as.data.frame() %>%
  select(-c("agy", "name")) %>% 
  rename(agy = agy_name, 
         name = job_group)


## grabbing summary trends for USDA and DOI: 
usda_doi <- enviro_tidy %>%
  mutate(name = str_to_title(name)) %>%
  mutate(name = gsub("U.s. ", "", name)) %>%
  mutate(agy = case_when(agy == "DOA" ~ "USDA", 
                         agy == "EPA" ~ "EPA", 
                         agy == "USACE" ~ "DoArmy", 
                         agy == "DOC-NOAA" ~ "DOC", 
                         TRUE ~ agy)) %>%
  mutate(agy_name = paste0(agy,  "-", name)) %>%
  mutate(agy_name = case_when(agy_name == "DOC-National Oceanic And Atmospheric Administration" ~ "DOC-NOAA", 
                              TRUE ~ agy_name)) %>%
  filter(agy %in% c("USDA", "DOI"))

usda_doi_filt <- usda_doi %>%
  group_by(agy, name, year, job_type_specific) %>%
  summarize(job_specific_employment = sum(employment)) %>%
  pivot_wider(names_from = job_type_specific, values_from = job_specific_employment) %>%
  mutate(STEM = STEM_analyst + STEM_research, 
         Tech = STEM_tech, 
         `Not STEM` = NOT_STEM) %>%
  select(agy, name, year, STEM:`Not STEM`) %>%
  group_by(agy, year) %>%
  summarize(STEM = sum(STEM, na.rm = T), 
            Tech = sum(Tech, na.rm = T), 
            `Not STEM` = sum(`Not STEM`)) %>%
  mutate(total_employees = STEM + Tech + `Not STEM`) %>%
  pivot_longer(STEM:`Not STEM`) %>%
  mutate(percent_workforce = (value/total_employees)*100)

stem_jobs_bureau <- stem_jobs_bureau[, names(usda_doi_filt)]

# fixing DoArmy :
stem_jobs_bureau <- stem_jobs_bureau %>%
  mutate(agy = case_when(agy == "DoArmy-Army Corps Of Engineers" ~ "Army Corps of Engineers", 
                         TRUE ~ agy))

# grabbing useful vectors and colors for plotting: 
doi <- c("DOI", "DOI-Bureau Of Land Management", "DOI-Bureau Of Reclamation", 
         "DOI-Fish And Wildlife Service", "DOI-Geological Survey", "DOI-National Park Service")
doi_colors <- c("#fb8500", "#ffb703",
                "#d62828", "#b5838d", 
                "#e5989b", "#8a817c")

usda <- c("USDA", "USDA-Forest Service", "USDA-Natural Resources Conservation Service")
usda_colors <- c("#9a8c98", "#4a4e69", "#bcb8b1")

# creating each plot - starting with DOI:
plot_data <- stem_jobs_bureau %>% filter(name != "Not STEM") %>% 
  filter(agy %in% doi)

doi_plot <- ggplot(plot_data, 
                   aes(x = year, 
                       y = percent_workforce, 
                       color = agy)) +
  geom_point(size = 1) +
  geom_smooth(se = FALSE, size = 1) + 
  geom_smooth(data = usda_doi_filt %>% filter(name != "Not STEM") %>% filter(agy == "DOI"),
              aes(text = sprintf("agy: DOI")),
              lty = "dashed", se = FALSE, size = 1, color = "#8a817c", name = "DOI") +
  geom_point(data = usda_doi_filt %>% filter(name != "Not STEM") %>% filter(agy == "DOI"),
              aes(text = sprintf("agy: DOI")),
             size = 1, color = "#8a817c", name = "DOI") +
  facet_wrap(~name, scales = "free_y") + 
    scale_color_manual(values = doi_colors,
                     name = "") +
  theme_minimal() +
  theme(legend.position = "bottom", 
        axis.text.x = element_text(angle = 90)) + 
  guides(color = guide_legend(ncol=3)) + 
  scale_x_continuous(breaks = seq(min(stem_jobs_wide$year),
                                  max(stem_jobs_wide$year),
                                  1)) + 
  labs(x = "Year", y = "% of Workforce") + 
  theme(text=element_text(size=11))
doi_plot
doi_plot <- plotly::ggplotly(doi_plot)

# USDA: 
plot_data <- stem_jobs_bureau %>% filter(name != "Not STEM") %>%
  filter(agy %in% usda)
usda_plot <- ggplot(plot_data, 
                   aes(x = year, 
                       y = percent_workforce, 
                       color = agy)) +
  geom_point(size = 1) +
  geom_smooth(se = FALSE, size = 1) + 
  geom_smooth(data = usda_doi_filt %>% filter(name != "Not STEM") %>% filter(agy == "USDA"),
              aes(text = sprintf("agy: USDA")),
              lty = "dashed", se = FALSE, size = 1, color = "#bcb8b1") +
  geom_point(data = usda_doi_filt %>% filter(name != "Not STEM") %>% filter(agy == "USDA"),
              aes(text = sprintf("agy: USDA")),
             size = 1, color = "#bcb8b1") +
  facet_wrap(~name, scales = "free_y") + 
  scale_color_manual(values = usda_colors,
                     name = "") +
  theme_minimal() +
  theme(legend.position = "bottom", 
        axis.text.x = element_text(angle = 90)) + 
  guides(color = guide_legend(ncol=3)) + 
  scale_x_continuous(breaks = seq(min(stem_jobs_wide$year),
                                  max(stem_jobs_wide$year),
                                  1)) + 
  labs(x = "Year", y = "% of Workforce") + 
  theme(text=element_text(size=11))
usda_plot
usda_plot <- plotly::ggplotly(usda_plot)

# plotting all others:
plot_data <- stem_jobs_bureau %>% filter(name != "Not STEM") %>%
  filter(!(agy %in% c(doi, usda))) 

all_others <- ggplot(plot_data, 
                   aes(x = year, 
                       y = percent_workforce, 
                       color = agy)) +
  geom_point(size = 1) +
  geom_smooth(se = FALSE, size = 1) + 
  facet_wrap(~name, scales = "free_y") + 
  scale_color_manual(values = c("#606c38", "#2a9d8f",
                                 "#84a98c"),
                     name = "") +
  theme_minimal() +
  theme(legend.position = "bottom", 
        axis.text.x = element_text(angle = 90)) + 
  guides(color = guide_legend(ncol=3)) + 
  scale_x_continuous(breaks = seq(min(stem_jobs_wide$year),
                                  max(stem_jobs_wide$year),
                                  1)) + 
  labs(x = "Year", y = "% of Workforce") + 
  theme(text=element_text(size=11))
all_others
all_others <- plotly::ggplotly(all_others)

# combining everything into a single plot: 
final_stem_tech <- subplot(all_others, doi_plot, usda_plot, nrows = 3, 
                           margin = 0.05)

```

## Plot Three: Ratio of IT to non-IT staff: 
```{r}
# adding department-level trends: 
dept_ratio <- usda_doi_filt %>%
  select(-c(total_employees, percent_workforce)) %>%
  pivot_wider(names_from = name, values_from = value) %>%
  mutate(non_it = STEM + `Not STEM`, 
        ratio_tech = Tech/non_it)

## ratio of IT to non-IT staff: 
ratio_tech <- ggplot(stem_jobs_wide, 
                     aes(x  = year, y = ratio_tech, 
                         color = agy_name)) + 
  geom_smooth(se = FALSE, size = 1.5) + 
  geom_point(size = 1.5) + 
  theme_minimal() + 
  geom_hline(yintercept = 0, lty = "dashed", color = "grey") + 
  geom_smooth(data = dept_ratio, aes(x = year, y = ratio_tech, color = agy),
              lty = "dashed", se = FALSE, size = 1.5) +
  geom_point(data = dept_ratio, aes(x = year, y = ratio_tech, color = agy),
             size = 1.5) +
  theme(legend.position = "bottom", 
        axis.text.x = element_text(angle = 90)) + 
  scale_color_manual(values = c("#606c38", "#2a9d8f", "#8a817c",
                                "#fb8500", "#ffb703",
                                "#d62828", "#b5838d", 
                                "#e5989b", "#84a98c", 
                                "#bcb8b1","#9a8c98", "#4a4e69"), 
                     name = "") + 
  labs(x = "Year", y = "IT staff / Non-IT staff") + 
  ggtitle("") + 
  theme(text=element_text(size=11))  + 
    guides(color=guide_legend(ncol = 3)) + 
  scale_x_continuous(breaks = seq(min(stem_jobs_wide$year),
                                  max(stem_jobs_wide$year),
                                  1)) 
ratio_tech
ratio_tech <- plotly::ggplotly(ratio_tech) 
```

## Creating HTML widgets: 
```{r}

htmlwidgets::saveWidget(partial_bundle(total_emp_plot) %>%
                          config(displayModeBar = FALSE) %>%
                          config(displaylogo = FALSE) %>%
                          layout(xaxis = list(fixedrange = TRUE), 
                                 yaxis = list(fixedrange = TRUE))%>%
                          layout(plot_bgcolor='transparent') %>%
                          layout(paper_bgcolor='transparent'),
                        "results/total_emp_interactive.html")

put_object(
   file = file.path("results/total_emp_interactive.html"),
   object = "/tech-capacity/total_emp_interactive.html",
   bucket = "tech-team-data"
)


htmlwidgets::saveWidget(partial_bundle(final_stem_tech) %>%
                          config(displayModeBar = FALSE) %>%
                          config(displaylogo = FALSE) %>%
                          layout(xaxis = list(fixedrange = TRUE), 
                                 yaxis = list(fixedrange = TRUE))%>%
                          layout(plot_bgcolor='transparent') %>%
                          layout(paper_bgcolor='transparent'),
                        "results/stem_tech_interactive.html")

put_object(
   file = file.path("results/stem_tech_interactive.html"),
   object = "/tech-capacity/stem_tech_interactive.html",
   bucket = "tech-team-data"
)


htmlwidgets::saveWidget(partial_bundle(ratio_tech) %>%
                          config(displayModeBar = FALSE) %>%
                          config(displaylogo = FALSE) %>%
                          layout(xaxis = list(fixedrange = TRUE), 
                                 yaxis = list(fixedrange = TRUE))%>%
                          layout(plot_bgcolor='transparent') %>%
                          layout(paper_bgcolor='transparent'),
                        "results/ratio_tech_interactive.html")

put_object(
   file = file.path("results/ratio_tech_interactive.html"),
   object = "/tech-capacity/ratio_tech_interactive.html",
   bucket = "tech-team-data"
)

```


## Extra figures for exploring other trends at agencies: 
```{r}
# Reed would like the interactive graphs! 
# Add these trends to ratio of IT staff to non-IT staff
# Departmental-level average, and then a subagency level? 

# Leader standpoint: 
# DOE, DOT, DOL, Department of Education, NIH, 
# DHS (similar in structure to DOI but the bureaus have a lot of autonomy since they were thrown together in the 2000s - CVP, FEMA, etc.)
# Make a subplot but make the scale comparable - keep the same y-axis 
# Hoping for this by Monday 22nd 

codes <- as.data.frame(unique(cube_data_test$agyt))
extra_agy_codes <- c("ED-DEPARTMENT OF EDUCATION", 
                     "DL-DEPARTMENT OF LABOR", 
                     "HE-DEPARTMENT OF HEALTH AND HUMAN SERVICES", 
                     "HS-DEPARTMENT OF HOMELAND SECURITY",
                     "TD-DEPARTMENT OF TRANSPORTATION", 
                     "NN-NATIONAL AERONAUTICS AND SPACE ADMINISTRATION",
                     "DN-DEPARTMENT OF ENERGY")

# filtering for agencies of interest:
extra_agy <- cube_data_test %>%
  filter(agyt %in% extra_agy_codes)

# in stem: 2210, 1550, 0854, 1515
tech_job <- c("1550", "2210", "1515", "0854")
# computer scientists, IT technology management, operations research, 
# computer engineering

# in stem: "1530" "1529" "1560" "0150" "1370"
data_analyst <- c("1560", "1370", "0150", "1530", "1529")
# data scientists, cartographers, geographers, cartographic technician, 
# geodetic technician, statistics, mathematical statistics 


# Grabbing stem jobs: 
extra_tidy <- extra_agy %>%
  mutate(job_type_broad = case_when(!(stemocc %in% c("XXXX", "****")) ~ "STEM_research", 
                                    TRUE ~ "NOT_STEM")) %>%
  mutate(job_type_specific = case_when(stemocc %in% tech_job ~ "STEM_tech",
                                       stemocc %in% data_analyst ~ "STEM_analyst", 
                                       TRUE ~ job_type_broad)) %>%
  mutate(job_type_broad = case_when(job_type_broad == "STEM_research" ~ "STEM", 
                                    TRUE ~ job_type_broad))

# grabbing percentage of staff in each category: 
extra_per <- extra_tidy %>%
  mutate(employment = as.numeric(employment)) %>%
  group_by(agyt, year, job_type_specific) %>%
  summarize(job_specific_employment = sum(employment)) %>%
  pivot_wider(names_from = job_type_specific, values_from = job_specific_employment) %>%
  mutate(STEM = STEM_analyst + STEM_research, 
         Tech = STEM_tech, 
         `Not STEM` = NOT_STEM) %>%
  select(agyt, year, STEM:`Not STEM`) %>%
  group_by(agyt, year) %>%
  summarize(STEM = sum(STEM, na.rm = T), 
            Tech = sum(Tech, na.rm = T), 
            `Not STEM` = sum(`Not STEM`)) %>%
  mutate(total_employees = STEM + Tech + `Not STEM`) %>%
  pivot_longer(STEM:`Not STEM`) %>%
  mutate(percent_workforce = (value/total_employees)*100)

# grabbing ratio of it to non-IT staff: 
extra_ratio <- extra_per %>% 
  select(-c(total_employees, percent_workforce)) %>%
  pivot_wider(names_from = name, values_from = value) %>%
  mutate(non_it = STEM + `Not STEM`, 
        ratio_tech = (Tech/non_it)) %>%
  mutate(year = as.numeric(year)) %>%
  mutate(agyt = str_to_title(agyt), 
         agyt = gsub(".*-","", agyt))


# plotting: 
extra_ratio_plot <- ggplot(extra_ratio, aes(x  = year, y = ratio_tech, 
                        color = agyt)) + 
  geom_smooth(se = FALSE, linewidth = 2) + 
  geom_point(size = 1.5) + 
  theme_minimal() + 
  geom_hline(yintercept = 0, lty = "dashed", color = "grey") + 
  theme(legend.position = "bottom", 
        axis.text.x = element_text(angle = 90)) + 
  labs(x = "Year", y = "IT staff / Non-IT staff") + 
  theme(text=element_text(size=11))  + 
  guides(color=guide_legend(ncol = 3)) + 
  scale_x_continuous(breaks = seq(min(extra_ratio$year),
                                  max(extra_ratio$year),
                                  1)) + 
  scale_color_discrete(name = "") + 
  ggtitle("Department Ratios of IT to Non-IT Staff")

extra_ratio_plot <- plotly::ggplotly(extra_ratio_plot) 


# also wanting to grab ratios for specific subagencies at DHS
dhs_sub <- extra_agy %>%
  filter(agyt == "HS-DEPARTMENT OF HOMELAND SECURITY") %>%
  mutate(agysubt = case_when(agysubt == "HSBE-FEDERAL LAW ENFORCEMENT TRAINING CENTER" ~ 
                               "HSBE-FEDERAL LAW ENFORCEMENT TRAINING CENTERS", 
         TRUE ~ agysubt)) %>%
  mutate(job_type_broad = case_when(!(stemocc %in% c("XXXX", "****")) ~ "STEM_research", 
                                    TRUE ~ "NOT_STEM")) %>%
  mutate(job_type_specific = case_when(stemocc %in% tech_job ~ "STEM_tech",
                                       stemocc %in% data_analyst ~ "STEM_analyst", 
                                       TRUE ~ job_type_broad)) %>%
  mutate(job_type_broad = case_when(job_type_broad == "STEM_research" ~ "STEM", 
                                    TRUE ~ job_type_broad)) %>%
  mutate(employment = as.numeric(employment)) %>%
  group_by(agysubt, year, job_type_specific) %>%
  summarize(job_specific_employment = sum(employment)) %>%
  pivot_wider(names_from = job_type_specific, values_from = job_specific_employment) %>%
  mutate(STEM = STEM_analyst + STEM_research, 
         Tech = STEM_tech, 
         `Not STEM` = NOT_STEM) %>%
  select(agysubt, year, STEM:`Not STEM`) %>%
  group_by(agysubt, year) %>%
  summarize(STEM = sum(STEM, na.rm = T), 
            Tech = sum(Tech, na.rm = T), 
            `Not STEM` = sum(`Not STEM`)) %>%
  mutate(total_employees = STEM + Tech + `Not STEM`) %>%
  pivot_longer(STEM:`Not STEM`) %>%
  mutate(percent_workforce = (value/total_employees)*100)

# grabbing ratio of it to non-IT staff: 
dhs_ratio <- dhs_sub %>% 
  select(-c(total_employees, percent_workforce)) %>%
  pivot_wider(names_from = name, values_from = value) %>%
  mutate(non_it = STEM + `Not STEM`, 
        ratio_tech = (Tech/non_it)) %>%
  mutate(year = as.numeric(year)) %>%
  mutate(agysubt = str_to_title(agysubt), 
         agysubt = gsub(".*-","", agysubt), 
         agysubt = gsub("U.s.","U.S.", agysubt), 
         agysubt = gsub("Dhs","DHS", agysubt))

all_dhs <-extra_ratio %>% filter(agyt == "Department Of Homeland Security")

# plotting: 
dhs_ratio_plot <- ggplot(dhs_ratio, aes(x = year, y = ratio_tech, 
                      color = agysubt)) + 
  geom_smooth(se = FALSE, linewidth = 2) + 
  geom_point(size = 1.5) + 
  theme_minimal() + 
  geom_hline(yintercept = 0, lty = "dashed", color = "grey") + 
  theme(legend.position = "bottom", 
        axis.text.x = element_text(angle = 90)) + 
  labs(x = "Year", y = "IT staff / Non-IT staff") + 
  theme(text=element_text(size=11))  + 
  guides(color=guide_legend(ncol = 3)) + 
  scale_color_discrete(name = "") + 
  scale_x_continuous(breaks = seq(min(extra_ratio$year),
                                  max(extra_ratio$year),
                                  1)) +
  geom_smooth(data = all_dhs, color = "black",
              se = FALSE, lty = "dashed") + 
  ggtitle("Department of Homeland Security - Subagency Ratios of IT to Non-IT Staff")


dhs_ratio_plot <- plotly::ggplotly(dhs_ratio_plot) 




# also wanting to grab ratios for specific subagencies at DHHS
dhhs_sub <- extra_agy %>%
  filter(agyt == "HE-DEPARTMENT OF HEALTH AND HUMAN SERVICES") %>%
  mutate(job_type_broad = case_when(!(stemocc %in% c("XXXX", "****")) ~ "STEM_research", 
                                    TRUE ~ "NOT_STEM")) %>%
  mutate(job_type_specific = case_when(stemocc %in% tech_job ~ "STEM_tech",
                                       stemocc %in% data_analyst ~ "STEM_analyst", 
                                       TRUE ~ job_type_broad)) %>%
  mutate(job_type_broad = case_when(job_type_broad == "STEM_research" ~ "STEM", 
                                    TRUE ~ job_type_broad)) %>%
  mutate(employment = as.numeric(employment)) %>%
  group_by(agysubt, year, job_type_specific) %>%
  summarize(job_specific_employment = sum(employment)) %>%
  pivot_wider(names_from = job_type_specific, values_from = job_specific_employment) %>%
  mutate(STEM = STEM_analyst + STEM_research, 
         Tech = STEM_tech, 
         `Not STEM` = NOT_STEM) %>%
  select(agysubt, year, STEM:`Not STEM`) %>%
  group_by(agysubt, year) %>%
  summarize(STEM = sum(STEM, na.rm = T), 
            Tech = sum(Tech, na.rm = T), 
            `Not STEM` = sum(`Not STEM`)) %>%
  mutate(total_employees = STEM + Tech + `Not STEM`) %>%
  pivot_longer(STEM:`Not STEM`) %>%
  mutate(percent_workforce = (value/total_employees)*100)

# grabbing ratio of it to non-IT staff: 
dhhs_ratio <- dhhs_sub %>% 
  select(-c(total_employees, percent_workforce)) %>%
  pivot_wider(names_from = name, values_from = value) %>%
  mutate(non_it = STEM + `Not STEM`, 
        ratio_tech = (Tech/non_it)) %>%
  mutate(year = as.numeric(year)) %>%
  mutate(agysubt = str_to_title(agysubt), 
         agysubt = gsub(".*-","", agysubt), 
         agysubt = gsub("U.s.","U.S.", agysubt))

all_dhhs <-extra_ratio %>% filter(agyt == "Department Of Health And Human Services")

# plotting: 
dhhs_ratio_plot <- ggplot(dhhs_ratio, aes(x = year, y = ratio_tech, 
                      color = agysubt)) + 
  geom_smooth(se = FALSE, linewidth = 2) + 
  geom_point(size = 1.5) + 
  theme_minimal() + 
  geom_hline(yintercept = 0, lty = "dashed", color = "grey") + 
  theme(legend.position = "bottom", 
        axis.text.x = element_text(angle = 90)) + 
  labs(x = "Year", y = "IT staff / Non-IT staff") + 
  theme(text=element_text(size=11))  + 
  guides(color=guide_legend(ncol = 3)) + 
  scale_color_discrete(name = "") + 
  scale_x_continuous(breaks = seq(min(extra_ratio$year),
                                  max(extra_ratio$year),
                                  1)) +
  geom_smooth(data = all_dhhs, color = "black",
              se = FALSE, lty = "dashed") + 
  ggtitle("Department of Health and Human Services - Subagency Ratios of IT to Non-IT Staff")


dhhs_ratio_plot <- plotly::ggplotly(dhhs_ratio_plot) 

```

## Exporting data for a spreadsheet
```{r}
dhhs_exp <- dhhs_ratio %>%
  janitor::clean_names() %>%
  mutate(agy = "DHHS") %>%
  relocate(agy)
dhs_exp <- dhs_ratio %>%
  janitor::clean_names() %>%
  mutate(agy = "DHS") %>%
  relocate(agy)
extra_exp <- extra_ratio %>%
  janitor::clean_names()

# stem, tech, not_stem, non_it, ratio_tech
stem_exp <- stem_jobs_wide%>%
  janitor::clean_names() %>%
  mutate(stem = stem_analyst + stem_research) %>%
  rename(tech = stem_tech, 
         non_it = non_tech) %>%
  select(agy, name, agy_name, year, stem, tech, not_stem, non_it, ratio_tech)
write.csv(stem_exp, "./results/csv/original_subagencies.csv")

head(dhs_exp)
write.csv(dhs_exp, "./results/csv/dhs.csv")
write.csv(dhhs_exp, "./results/csv/dhhs.csv")

# agency ratios: 
write.csv(extra_exp, "./results/csv/extra_agy.csv")

dept_ratio_exp <- dept_ratio %>%
  janitor::clean_names() %>%
  mutate(agy = case_when(
    agy == "DOI" ~ "Department Of Interior", 
    agy == "USDA" ~ "U.S. Department of Agriculture"
  ))
write.csv(dept_ratio_exp, "./results/csv/original_agencies.csv")

```


